\Section{Related Work}
\label{sec:related}

Scratch is a tile-based scripting language for kids. A pervasive simplicity and a well-polished user interface make Scratch a welcoming environment for first-time users.
Naturally, this ease of use comes at a price.  Unlike most other programming languages, Scratch
provides no form of abstraction (e.g., users cannot create their own tiles or functions.) Scratch projects that go beyond a certain threshold of
sophistication tend to contain unwieldy programs that can be difficult to write and to maintain.

%Scratch is a tile-based scripting language for kids. Unlike most other programming languages, Scratch
%does not support any form of abstraction (e.g., users cannot create their own tiles or functions); this level of
%simplicity, combined with a well-polished user interface, makes Scratch a welcoming environment for first-time users.
%Unfortunately, this ease of use comes at a high price: Scratch projects that go beyond a certain threshold of
%sophistication tend to contain unwieldy programs that are frustrating to write and to maintain.

Our own Squeak Etoys, another easy-to-learn tile-based scripting language, supports user-defined
procedures, or {\em scripts}. Although Etoys scripts cannot return values (which makes them less powerful than
functions), they provide an important form of abstraction that in turn enables users to tackle moderately
sophisticated projects without ``feeling like a dog standing on its hind legs''. This is clearly a step in
the right direction, but we can do better.

TileScript is our attempt to create an end-user programming language that is just as welcoming to new programmers as
Scratch and Etoys, but which scales to arbitrarily complex tasks like a conventional programming language.

Squeak's Universal Tiles (UniTiles) was an earlier attempt at creating a more scalable end-user scripting system.
UniTiles was isomorphic to Smalltalk, which made it just as expressive as---but no easier to
learn than---Smalltalk. Additionally, because Smalltalk does not support macros, a UniTiles implementation of our {\tt
repeat} example would require its {\tt body} argument to be passed in as a block, which can be confusing to beginners.
TileScript's macro construct and its associated user interface component make it easy for end-users to define their own
control structures.

Like TileScript, IMP~\cite{i70imp} allowed programmers to define their own control structures using a macro-like mechanism.
The main difference between these two approaches is that IMP was based on syntax extension: each new control structure
was accompanied by a BNF description of its intended syntax. The analog of syntax extension in TileScript is our
tile-creation GUI, which allows end-users to design the look-and-feel of the tile associated with a particular macro. We
believe this approach to provide a better fit with our intended users, who are not expert programmers.
